\LARGE \bf A Formal Description of Bitcoin's Script Language

### 2017-03-20 paper.tex
Removed 3rd application, won't do it.
rough rough ugly draft of 'Approach'

% The following packages can be found on http:\\www.ctan.org
%\usepackage{graphics} % for pdf, bitmapped graphics files
%\usepackage{epsfig} % for postscript graphics files
%\usepackage{mathptmx} % assumes new font selection scheme installed
%\usepackage{times} % assumes new font selection scheme installed
\usepackage{amsmath} % assumes amsmath package installed
\usepackage{amssymb}  % assumes amsmath package installed
\usepackage{xcolor}

@@

\subsection{Definition of Script Grammar}

\subsubsection{Syntax}
\# Bitcoin wiki -> tables describe the format\\
\# transaction is X Y, inputs are such and such\\
\# looked at enums in source code to get full list of defined opcode and their hex values\\
JOSEPH: also explain how the different syntactic groups were broken down\\

@@

We described the big-step semantics and the small-step semantics of the language using a combination of the reference implementation of the language [Ref] and of the Script page of the Bitcoin wiki [Ref].

@@

@@

\subsubsection{Background}
For the Bitcoin network to function properly, it is important that the network form a single connected component. A type of attack against the network is to try and cut some nodes from the rest of the network, and then ``double-spend'' the same amount of money on the two disconnected components, which will both accept the transactions. Such a cut can be achieved by performing a DoS attack against a group of nodes that collectively ensure connectivity between two otherwise separated subnetworks. One way to perform a DoS attack could be to have a node attempt to evaluate a Script program with an unreasonable execution time. It is therefore important to guarantee that a Script program will terminate within a certain amount of steps.

\subsubsection{Analysis}
Using the small-step semantics defined in Section 3 of the appendix, we see that no reduction rule yields its entire input, as was the case with the \texttt{while} local reduction rule in IMP. That is to say, no recursion is allowed. At most a proper subset of the input is conserved after applying the rule, when it is not completely replaced by \texttt{OP\_NOP}, Script's equivalent of the \texttt{skip} command. This guarantees that every small-step reduction progresses towards termination of the program. Furthermore, the global reduction rule trivially entails that for any program that is not only \texttt{OP\_NOP}, there exists a unique context reducible by a local reduction rule. Thus we have that every Script program halts.

@@

\subsection{Previous versions of Script allowed skeleton key scripts}

\subsubsection{Background}
The way an amount on Bitcoin is spent on the network is by evaluating to $\sigma[V=valid]$\footnote{See Appendix section 2 for a description of state configurations.} the concatenation $C = BA$ of a locking script A, provided by the payer, and an unlocking script B provided by the recipient. The intended scenario is that $A$ will contain a validation condition, which will check whether the output of $B$ is adequate. Typically, $B$ will have to provide a hash that proves they are the intended recipient of the money spent by the payer, and $A$ will know what to expect as a proof. In that way, $A$ is a guarantee to the payer that the money can only be received by the intended recipient. However, a careless definition of the \texttt{OP\_RETURN} opcode in previous versions of Script made it possible to program skeleton key scripts that could spend any amount of money, without proving the recipient's identity.

\subsubsection{Analysis}
A skeleton script would have the following property: 
$$\exists B. \forall A. \hspace{1.5mm} BA \Downarrow \sigma[V=valid]$$

This is equivalent to saying that either (1) $B$ is such that for any $A$, $BA$ evaluates to a valid state, or (2) $B$ evaluates to a state from which any $A$ will evaluate to a valid state. Case (2) was never possible, since one could write a script $A = \texttt{OP\_0}$, that pushes 0 on the stack. Upon normal program termination, if the top stack value is 0, then the program is deemed invalid. Thus for this $A$ there is no script $B$ such that $BA \Downarrow \sigma[V=valid]$. Case (1), however, was made possible by the semantics of \texttt{OP\_RETURN} prior to July 31 2010. In local reduction rule form, \texttt{OP\_RETURN} was defined as follows:
%
$$<\texttt{OP\_RETURN com}, \sigma> \hspace{3mm} \longrightarrow \hspace{3mm} <\texttt{OP\_NOP}, \sigma>$$
%
Evaluating \texttt{OP\_RETURN} results in the immediate termination of the program, \textit{without any changes to the state}. Consider now a script $B = \texttt{OP\_1 OP\_RETURN}$, and the evaluation of $BA$ for any arbitrary script $A$:

\begin{enumerate}
  \item \texttt{1} is pushed onto the stack
  \item \texttt{OP\_RETURN} triggers program termination
  \item The top stack value is nonzero, and so the program is deemed valid
\end{enumerate}

Clearly, that such a script was possible completely defeated the transaction validation process. \texttt{OP\_RETURN} semantics were subsequently changed to the following:
%
$$<\texttt{OP\_RETURN com}, \sigma> \hspace{1mm} \longrightarrow \hspace{1mm} <\texttt{OP\_NOP}, \sigma[V=invalid]>$$
%
thus ensuring that its evaluation in a script invariably resulted in a rejected transaction.
As no other command in the language can cause early termination of the program with a valid state, we can now say that there exists no skeleton key script that could serve to unlock any arbitrary locking script.

\subsection{The Script command set could be reduced without loss of expressiveness}
Here we prove that some commands are redundant. Must explain why this can be desirable sometimes, but also why some of them are clearly uselessly redundant.

>>>>

% The following packages can be found on http:\\www.ctan.org
%\usepackage{graphics} % for pdf, bitmapped graphics files
%\usepackage{epsfig} % for postscript graphics files
%\usepackage{mathptmx} % assumes new font selection scheme installed
%\usepackage{times} % assumes new font selection scheme installed
%\usepackage{amsmath} % assumes amsmath package installed
%\usepackage{amssymb}  % assumes amsmath package installed
\usepackage{xcolor}

@@

\subsection{Definition of Script Grammar}
The bulk of our project consists in providing a well-formed and consistent definition of the Script language. This includes a description of the syntax that the text of scripts must conform to, as well descriptions of the operational semantics--big-step and small-step--by which the scripts evaluate. Due to its breadth, we defer the full description of this grammar to [Appendix A].

The reference implementation of Bitcoin defines the network protocol. Script, as the language which expresses and encodes transactions, exists as a subset of the Bitcoin protocol and is therefore also defined by the reference implementation.

In order to generate this grammar, it was ultimately necessary to directly turn to the source code of the reference implementation to validate the correctness of our grammar. Proving the consistency or equivalence of our work with the reference implementation is outside the scope of this project and is left as future work. In the process of forming both our understanding and therefrom

we searched through several reputable sources in order to gain a working understanding of Script, eventually turning to the reference Bitcoin implementation to secure and progressively refine our understanding.

\subsubsection{Syntax}

First, it was necessary to define the syntax. We made heavy use of the [Bitcoin Wiki] to this end. Script is a Forth-like, stack-based language. It is parsed and evaluated left to right [wiki/script]. As it is defined by it's reference implementation, in order to implement the language one must ensure that in all circumstances one's implementation interprets the binary serialization of a script in a manner compatible with the reference.

It was therefore quite natural to define the syntax of the literal bytes that form programs. This is consistent with the internal subdivisions one may discern within a script program. Script defines words or opcodes with defined effects on the state of the program. We look to both the [Bitcoin Wiki] and the source code to retrieve a full list of defined opcodes and to infer and ultimately describe their operations.

The syntax in [Appendix A] defines terms the form of a transaction, with scripts existing within transactions in defined positions. We provide aliases for opcodes' hex values and group commands by both their effect--arithmetic, control, crypto, etc, and their form--whether they are "atomic" or whether the opcode relies on following words or bytes to affect how it transform the state of the program and to give the opcode syntactic validity--the script \texttt{OP\_IF OP\_0 OP\_ELSE} is invalid because it has no \texttt{OP\_ENDIF}.

@@

After providing a sufficient syntax, we moved to describe the operational semantics. We described the big-step semantics and the small-step semantics of the language using a combination of the reference implementation of the language [Ref] and of the Script page of the Bitcoin wiki [Ref].

We provide inference rules for every opcode. The state of the program consists of the stack, $S$, the alt-stack, $AS$, the validity $V$, and the rest of the program to be parsed $P$. Script operates primarily on byte-vectors representing numbers or arbitrary data. We speak of a byte-vector evaluating to a number under a set of encoding rules specific to Bitcoin but not completely described.

A number of op-codes have been disabled due to perceived vulnerabilities, reserved for future use, or reserved for internal use in testing. These are given their own syntactic category and result in an invalid state.

Scripts exist in transactions and op-codes exist which directly reference their transaction context. We omit these because of the disproportionate descriptive burden they introduce.

We also provide small step semantics to describe the progression of evaluation of a script program. The local and global reduction rules are compact due to the language's inherent simplicity.

@@

 \textcolor{red}{\textit{If appropriate, you should describe your implementation and experimental results. Explain how to interpret your numbers and results if applicable.}}

\textcolor{blue}{[Motivation for the application cases we chose.]}

@@

For the Bitcoin network to function properly, it is important that the network form a single connected component. A type of attack against the network is to try and cut some nodes from the rest of the network, and then ``double-spend'' the same amount of money on the two disconnected components, which will both accept the transactions. Such a cut can be achieved by performing a DoS attack against a group of nodes that collectively ensure connectivity between two otherwise separated subnetworks. One way to perform a DoS attack could be to have a node attempt to evaluate a Script program with an unreasonable execution time. It is therefore important to guarantee that a Script program will terminate within a certain amount of steps.

Using the small-step semantics defined in Section 3 of the appendix, we see that no reduction rule yields its entire input, as was the case with the \texttt{while} local reduction rule in IMP. That is to say, no recursion is allowed. At most a proper subset of the input is conserved after applying the rule, when it is not completely replaced by \texttt{OP\_NOP}, Script's equivalent of the \texttt{skip} command. This guarantees that every small-step reduction progresses towards termination of the program. Furthermore, the global reduction rule trivially entails that for any program that is not only \texttt{OP\_NOP}, there exists a unique context reducible by a local reduction rule. Thus we have that every Script program halts.

@@

\subsection{Previous versions of Script allowed undesirable programs}
Here we prove how previous states were problematic

\subsection{The Script command set could be reduced without loss of expressiveness}
Here we prove that some commands are redundant. Must exaplin why this can be desirable sometimes, but also why some of them are clearly uselessly redundant.
